Testing atlantisom: allow users to specify different max length bin for length compositions

### 2019-05-15 TestBiggerMaxLen.Rmd
testing changes to calc_age2length

---
title: "Testing atlantisom: generate survey sampled length comps for Norwegian Barents Sea Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

@@

## Simulate a survey part 4: sample for length composition

@@

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.

```{r stdsurveyNbased, echo=TRUE}

survey_testNstd <- create_survey(dat = truth$nums,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv,
                                 effic = efficmix,
                                 selex = selexmix)

# consider saving this interim step if it takes a long time go generate

Next we apply the aggregateDensityData function can to resn and structn for survey times, species, and boxes.

```{r aggdens, echo=TRUE}

# aggregate true resn per survey design
survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv)

# aggregate true structn per survey design
survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv)

Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey. The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

@@

# apply default sample fish as before to get numbers
numsstd <- sample_fish(survey_testNstd, effNmix)

#dont sample these, just aggregate them using median (effNmix does nothing)
structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)

resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)
```

```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE, echo=TRUE}

# now cut these down to a single species for testing

ss_numsstd <- numsstd[numsstd$species == "Green_halibut",]
ss_structnstd <- structnstd[structnstd$species == "Green_halibut",]
ss_resnstd <- resnstd[resnstd$species == "Green_halibut",]

ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
                                 resn = ss_resnstd,
                                 nums = ss_numsstd,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

Plot samples one species:

```{r sslengthsamp1-test}

lfplot <- ggplot(ss_length_stdsurv$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_stdsurv$natlength$species))

# try dir = "v" option for vertical lf comparisons

@@

Sampling looks really different for greenland halibut than census did. Check this!

## Samples for all species and SAVE the file

Now run for all species in the survey.

```{r allspplengthsamp, echo=TRUE, eval=FALSE}

length_stdsurv <- calc_age2length(structn = structnstd,
                                 resn = resnstd,
                                 nums = numsstd,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

#save for later use, takes a long time to generate
saveRDS(length_stdsurv, file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))

@@

Runtime for census lengths was ~2.5 hours for NOBA once we decided once per year was plenty. For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.

Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html):

```{r loadlengthcomp, echo=TRUE}

length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

length_stdsurv <- readRDS(file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))

@@

Comparison plots for other species. Blue bars are the census length sample and red outlines are the survey sample (standardized for comparison; census total n lengths is 1e+8 while sample total n lengths is 1000):

```{r moreplots}

# select a species
ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
  filter(species == "Norwegian_ssh")

ss_length_stdsurv <- length_stdsurv$natlength %>%
  filter(species == "Norwegian_ssh")

# make proportion at length to plot together
lf_census_tot <- aggregate(ss_length_censussurvsamp$atoutput,list(ss_length_censussurvsamp$time),sum )
names(lf_census_tot) <- c("time","totlen")

lf_census_prop <- merge(ss_length_censussurvsamp, lf_census_tot)

lf_samp_tot <- aggregate(ss_length_stdsurv$atoutput,list(ss_length_stdsurv$time),sum )
names(lf_samp_tot) <- c("time","totsamp")

lf_samp_prop <- merge(ss_length_stdsurv, lf_samp_tot)

# add sample, this is census, probably want it first?
lfplot <- ggplot(mapping=aes(x=upper.bins)) +
  geom_bar(data=lf_census_prop, aes(weight = atoutput/totlen), fill="blue") +
  geom_bar(data=lf_samp_prop, aes(weight = atoutput/totsamp), colour="red") +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, lf_census_prop$species))

@@

Here as with [cohort sampling](https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html), the impact of survey selectivity is much larger than sampling error with effN of 1000. We could try effN that is lower, but 1000 lengths may be reasonable from a survey. A `sample_lengths` function could add further measuremet error on top of this, if desired (though measurement error at the 1 cm level may be negligible).

>>>>

---
title: "Testing atlantisom: allow users to specify different max length bin for length compositions"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

Here we briefly test changes to the function `calc_age2length` that should allow the user to specify a different max length bin. For all setup, etc, please see previous files Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

@@

## Simulate a survey part 4: sample for length composition (testing revised function)

@@

We will apply examples here to only one species, Greenland halibut in NOBA, which grows to a large size.

@@

And the numbers to be sampled for lengths each year:

@@

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

```{r stdsurveyNbased-GHR, echo=TRUE}

ss.name <- funct.group.names[funct.group.names == "Green_halibut"]

#change back to flat selectivity to see full comp
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# get survey nums with full (no) selectivity
ss_survey_testNstd_nosel <- create_survey(dat = truth$nums,
                                 time = annualmidyear,
                                 species = ss.name,
                                 boxes = boxsurv,
                                 effic = efficmix,
                                 selex = selex1)

# now sample fish nums from this
ss_numsstd_nosel <- sample_fish(ss_survey_testNstd_nosel, effNmix)

# aggregate true resn per survey design
survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
                                 time = annualmidyear,
                                 species = ss.name,
                                 boxes = boxsurv)

# aggregate true structn per survey design
survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
                                 time = annualmidyear,
                                 species = ss.name,
                                 boxes = boxsurv)

#dont sample these, just aggregate them using median (effNmix does nothing)
ss_structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)

ss_resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)

@@

Length sample with default max length bin (150 cm):

```{r default-maxlen, echo=TRUE}

ss_length_stdsurv_nosel <- calc_age2length(structn = ss_structnstd,
                                 resn = ss_resnstd,
                                 nums = ss_numsstd_nosel,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

Length sample with user specified max length bin (250 cm):

```{r userset-maxlen, echo=TRUE}

ss_length_stdsurv_nosel_max <- calc_age2length(structn = ss_structnstd,
                                 resn = ss_resnstd,
                                 nums = ss_numsstd_nosel,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 250,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

Plots show that default length bins are not adequate for something like Greenland halibut, which get larger than the default largest bin in `calc_age2length`, 150 cm. Both census and these sampled length comps are chopped off at 150 cm for this species:

```{r sslengthsamp2-testdefault}

lfplot <- ggplot(ss_length_stdsurv_nosel$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_stdsurv_nosel$natlength$species))

@@

I changed the `calc_age2length` code to pass the upper bin of 150 as a default but allow the user to change it in the function call. This plot shows results from setting maxbin = 250 for Greenland Halibut in NOBA:

```{r sslengthsamp2-testmax250}

lfplot <- ggplot(ss_length_stdsurv_nosel_max$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        ss_length_stdsurv_nosel_max$natlength$species))

Now we can go back and apply standard survey selectivity as in the [NOBA visualization](https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html) and see how the length comps look:

```{r stdsurveyNbased-GHR2, echo=TRUE}

ss.name <- funct.group.names[funct.group.names == "Green_halibut"]

# get survey nums with standard selectivity
ss_survey_testNstd <- create_survey(dat = truth$nums,
                                 time = annualmidyear,
                                 species = ss.name,
                                 boxes = boxsurv,
                                 effic = efficmix,
                                 selex = selexmix)

# now sample fish nums from this
ss_numsstd <- sample_fish(ss_survey_testNstd, effNmix)

# structn and resn stuff is exactly the same because selectivity is irrelevant
```

Length sample from standard survey with user specified max length bin (250 cm):

```{r user-maxlen2, echo=TRUE}

ss_length_stdsurv_max <- calc_age2length(structn = ss_structnstd,
                                 resn = ss_resnstd,
                                 nums = ss_numsstd,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 250,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

```{r sslengthsamp3-testmax250}

lfplot <- ggplot(ss_length_stdsurv_max$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        ss_length_stdsurv_max$natlength$species))

Bigger max bin is definitely more appropriate for this species.

I suppose we could make the maxbin a vector by species, but for now may be overkill.
